Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-09-18 2017/ParOptBook/Chapter4.tex
ParOptBook: update Chapter4.tex

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
%                           % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%\usepackage{mathtools}
%\mathtoolsset{showonlyrefs}
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

@@

As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$ re assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
\begin{equation}
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2
\end{equation}

@@

\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
\[
R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\]
\begin{equation}%\label{eq:14}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\[
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\]
where $\Delta_i=(x_i - x_{i-1})^{1/N}$. The values $r_\nu > 1, \; 1 \leq
\nu \leq m+1,$ are parameters of the algorithm. An appropriate selection
of $r_\nu$ allows to consider the product $r_\nu \mu_\nu$ as an estimate
of the H{\"o}lder constants $H_\nu, \; 1 \leq \nu \leq m+1$.

\emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}\label{MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}

Rule 7. Make the next trial at the midpoint of the interval
$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not
the same, i.e.
\[
x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\]
Otherwise, make the trial at the point
\begin{equation}%\label{eq:142}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{\mathrm{sign}(z_t-z_{t-1})}{2r_\nu}\left[\frac{\left|z_t-z_{t-1}\right|}{\mu_\nu}\right]^N, \; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}

We may take as termination condition the inequality $\Delta_t \leq
\epsilon$, where $t$ is from (\ref{MaxR}) and $\epsilon>0$ is the
predefined accuracy.

Various modifications of this algorithm and the corresponding theory of
convergence are given in
\cite{Strongin2000,Barkalov2002,Gergel2005,Strongin2013}.

The algorithm considered above is very flexible and allows for an
efficient parallelization for shared memory, for distributed memory, and
for accelerators \cite{Barkalov2010,Barkalov2015,Barkalov2016,Gergel2016}.

@@

%\end{document}

>>>>

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                           % not available on your system

\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

@@

As before, the objective function $\varphi$ (hereinafter denoted also as $g_{m+1}$) and the left parts of the constraints $g_j(x)\le 0,1\le j\le m$ re assumed to satisfy \emph{Lipschitz condition} with the corresponding constants $L_j,1\le j\le m+1$, namely:
\begin{equation}
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2\in [a,b]
\end{equation}

@@

\emph{Rule 6.} For each interval ($x_{i-1},x_i), \; 1 \leq i \leq k+1,$ compute the \textit{characteristics} $R(i)$:
\[
R(i)=2\Delta_i-4\frac{z_i-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_i)>\nu(x_{i-1}),
\]
\begin{equation}
\label{eq4:characteristic}
R(i)=\Delta_i+\frac{(z_i-z_{i-1})^2}{r_\nu^2 \mu_\nu^2\Delta_i}-2\frac{z_i+z_{i-1}-2z_\nu^\ast}{r_\nu \mu_\nu}, \;  \nu=\nu(x_i)=\nu(x_{i-1}),
\end{equation}
\[
R(i)=2\Delta_i-4\frac{z_{i-1}-z_\nu^\ast}{r_\nu \mu_\nu}, \; \nu=\nu(x_{i-1})>\nu(x_i),
\]
\[
\Delta_i=x_i - x_{i-1}
\]
The values $r_\nu>1, 1\le\nu\le m+1$ are the parameters of the algorithm. An appropriate choice of the values $r_\nu$ allows using the products $r_\nu\mu_\nu$ as the estimates of Lipschitz constants $L_\nu, 1\le\nu\le m+1$.
\emph{Rule 7.} Find the interval $(x_{t-1},x_t)$ with the maximum characteristic
\begin{equation}
\label{eq4:MaxR}
R(t)=\max{\left\{R(i): 1 \leq i \leq k+1\right\}}.
\end{equation}
\emph{Rule 8.} Make the next trial at the midpoint of the interval
$(x_{t-1},x_t)$ if the indices of the points $x_{t-1}$ and $x_t$  are not the same, i.e.
\[
x^{k+1} = \frac{x_t + x_{t-1}}{2}, \; \nu(x_{t-1}) \neq \nu(x_t).
\]
Otherwise, make the trial at the point
\begin{equation}
\label{eq4:next_point}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \frac{z_t-z_{t-1}}{2r_\nu\mu},\; \nu=\nu(x_{t-1})=\nu(x_t).
\end{equation}

The described rules can be supplemented with the termination condition, finalizing the iterations of the method if
\begin{equation}
\label{eq4:stop_cond}
  x_t - x_{t-1}\le\epsilon
\end{equation}
where t from \eqref{eq4:MaxR} and $\epsilon>0$ is the predefined accuracy of the optimization.

Let us formulate the conditions of convergence for the algorithm in the form of the following theorem.
\begin{theorem}
Let the considered index algorithm be applied to the solving of problem \eqref{eq4:problem} and the following conditions be satisfied:
\begin{enumerate}
  \item each function $gj, 1\le j\le m+1$ satisfies Lipschitz condition with the constant $L_j$ in the interval $[a,b]$, i.e.,
  \[
  |g_j(x_1)-g_j(x_2)|\le L_j|x_1-x_2|,1\le j\le m+1,x_1,x_2 \in[a,b]
  \]
  \item the following inequalities hold for the quantities $\mu_\nu$ from \eqref{eq4:lip_est} starting from some step
  \[
  \r_\nu\mu_\nu > 2L_\nu,1\le\nu\\le m+1.
  \]
\end{enumerate}

Then the set of the limit points of the sequence $\{x_k\}$ generated by the index algorithm coincides with the set of solutions of problem \eqref{eq4:problem} with $\epsilon=0$ and termination criterion \eqref{eq4:stop_cond}. In addition, the index of each limit point equals to $M$ and the convergence to any limit point $\overline{x}$  is bilateral if $\overline{x}\not=a$ and $\overline{x}\not=b$.
\end{theorem}

This theorem is given without a proof, which can be found, for example, in \cite{}[1, 2]. Various modifications of this algorithm and the corresponding theory of convergence are given in \cite{}[5–10].

@@

\end{document}
